---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(modelr)
library(broom)
```

Load the data into a dataframe project
```{r}
project_management <- read_csv("data/project_management.csv") %>% janitor::clean_names()
```


Plot the data, taking estimated_length as the independent variable and actual_length as the dependent variable.
```{r}
project_management %>% 
  ggplot(aes(x = estimated_length,
             y = actual_length))+
  geom_point()
```


Calculate the correlation coefficient of estimated_length and actual_length and interpret the value you obtain.
```{r}
project_management %>% 
  summarise(correlation = cor(estimated_length, actual_length))
```
This falls within the ranges of 0.80 - 0.99 which suggests a very strong correlation


Perform a simple linear regression using actual_length as the dependent variable, and estimated_length as the independent variable. Save the model object to a variable.
```{r}
model <- lm(actual_length ~ estimated_length,
            data = project_management)
```


Interpret the regression coefficient of estimated_length (i.e. slope, gradient) you obtain from the model. How do you interpret the r2 value reported by the model?
```{r}
glance_output <- model %>% 
  glance() %>% 
  janitor::clean_names()

glance_output
```
The r^2 value of 0.65 means that the model is an OK fit and therefore, the actual days for a job can only be estimated to an OK degree of certainty


Is the relationship statistically significant? Remember, to assess this you need to check the p-value of the regression coefficient (or slope/gradient). But you should first check the regression diagnostic plots to see if the p-value will be reliable (don’t worry about any outlier points you see in the diagnostic plots, we’ll return to them in the extension).
```{r}
library(ggfortify)

autoplot(model)
```
From the diagnostic plots we can see:

Residuals vs Fitted:
The residuals tend to descrease as the fitted values increase, which shouuldn't be the case and suggests that any p-value could be misleading

Scale-Location:
Again, there seems to be a pattern to the smoothed line which is undesirable and agains suggests that the p-value cannot be accurate

```{r}
glance_output
```
The p-value is tending towards 0 which is suggesting that there is a statistically significant relationship between the actual days and estimated days of a job



```{r}
plot(model) +
  geom_text(aes(label = 1:nrow(project_management)))
```



