---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(janitor)
library(infer)
```

```{r}
ames <- read_csv("data/ames.csv") %>% 
  clean_names()
```


```{r}
glimpse(ames)
```

# Task 2.
Investigate the distribution of lot_area. Is the distribution roughly normal? If not, what problems do you find?
```{r}
ames %>% 
  ggplot(aes(x = lot_area)) +
  geom_histogram(col = "white")
```
The plot is unimodal with a strong peak
It is uniformly distributed
There is a strong peak


# Task 3.
Compute and visualise a bootstrap sampling distribution for the `mean(lot_area)` of the sold houses.
```{r}
bootstrap_sample <- ames %>% 
  rep_sample_n(size = 2930,
               reps = 5000,
               replace = TRUE) %>% 
  summarise(mean_lot_area = mean(lot_area)
            )

bootstrap_sample
```


```{r}
bootstrap_sample %>% 
  ggplot(aes(x = mean_lot_area)) +
  geom_histogram(col = "white")
```


# Task 4.
Use your bootstrap distribution to calculate a 95% CI for mean(lot_area), and visualise it on the distribution.
```{r}
ci_95 <- bootstrap_sample %>% 
  summarise(
    mean = mean(mean_lot_area),
    lower_bound = quantile(mean_lot_area, probs = 0.05),
    upper_bound = quantile(mean_lot_area, probs = 0.95)
  )

ci_95
```


```{r}
infer_resample <- ames %>% 
  specify(response = lot_area) %>% 
  generate(reps = 5000, type = "bootstrap") %>% 
  calculate(stat = "mean")

infer_resample
```


```{r}
infer_ci_95 <- infer_resample %>% 
  get_confidence_interval(level = 0.95, type = "percentile")

infer_ci_95
```


```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_95)
```


# Task 5.
You would like to know the mean(lot_area) of the sold houses with higher confidence. Calculate the 99% CI for this variable (you can re-use your bootstrap distribution from above). Is it narrower or broader than the 95% CI? Does that make sense?
```{r}
infer_ci_99 <- infer_resample %>% 
  get_confidence_interval(level = 0.995, type = "percentile")

infer_ci_99
```


```{r}
infer_resample %>% 
  visualise(bins = 30) +
  shade_confidence_interval(endpoints = infer_ci_99)
```


# Task 6.
Calculate the point estimate of the mean(lot_area)
```{r}
infer_resample %>% 
  summarise(point_estimate = mean(stat))
```


