---
title: "R Notebook"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidytext)
library(janeaustenr)
library(textdata)
```

## Find the most common words in both Pride & Prejudice and Sense & Sensibility.

### Pride and Prejudice

```{r}
pride_and_prej <- tibble(
  id = 1:length(prideprejudice),
  text = prideprejudice
) %>% 
  unnest_tokens(word, text)

pride_and_prej
```


```{r}
pride_and_prej %>% 
  count(word, sort = TRUE)
```


### Sense & Sensibility

```{r}
sense <- tibble(
  id = 1:length(sensesensibility),
  text = sensesensibility
) %>% 
  unnest_tokens(word, text)

sense
```


```{r}
sense %>% 
  count(word, sort = TRUE)
```


## Find the most common words in both Pride & Prejudice and Sense & Sensibility, not including stop words.

### Pride and Prejudice
```{r}
pride_no_stop <- pride_and_prej %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE)

pride_no_stop
```


### Sense & Sensibility 
```{r}
sense_no_stop <- sense %>% 
  anti_join(stop_words) %>% 
  count(word, sort = TRUE)

sense_no_stop
```


## Find the most common sentiment words in both Pride & Prejudice and Sense & Sensibility.


### Pride and Prejudice
```{r}
pride_no_stop %>% 
  inner_join(get_sentiments("afinn")) %>% 
  count(word, sort = TRUE)

pride_no_stop %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(word, sort = TRUE)

pride_no_stop %>% 
  inner_join(get_sentiments("nrc")) %>% 
  count(word, sort = TRUE)

pride_no_stop %>% 
  inner_join(get_sentiments("loughran")) %>% 
  count(word, sort = TRUE)
```


### Sense & Sensibility 
```{r}
sense_no_stop %>% 
  inner_join(get_sentiments("afinn")) %>% 
  count(word, sort = TRUE)

sense_no_stop %>% 
  inner_join(get_sentiments("bing")) %>% 
  count(word, sort = TRUE)

sense_no_stop %>% 
  inner_join(get_sentiments("nrc")) %>% 
  count(word, sort = TRUE)

sense_no_stop %>% 
  inner_join(get_sentiments("loughran")) %>% 
  count(word, sort = TRUE)
```




