---
title: "R Notebook"
output: html_notebook
---


```{r}
library(tidyverse)
library(GGally)
library(janitor)
library(mosaic)
library(ggfortify)
```


Load the housing_prices.csv data set and undertake an initial exploration of the data. You will find details on the data set on the relevant Kaggle page
```{r}
houses <- read_csv("data/housing_prices.csv")
```


```{r}
houses %>% 
  summary()
```


We expect the total_rooms of houses to be strongly correlated with total_bedrooms. Use ggpairs() to investigate correlations between these two variables.
```{r}
houses %>% 
  select(total_rooms, total_bedrooms) %>% 
  ggpairs()
```

So, we do find significant correlations. Let’s drop total_bedrooms from the dataset, and use only total_rooms going forward.
```{r}
houses <- houses %>% 
  select(-total_bedrooms)
```
We are interested in developing a regression model for the median_house_value of a house in terms of the possible predictor variables in the dataset.

Use ggpairs() to investigate correlations between median_house_value and the predictors (this may take a while to run, don’t worry, make coffee or something).

```{r}
houses %>% 
  ggpairs()
```

Perform further ggplot visualisations of any significant correlations you find.

```{r}
houses %>% 
  ggplot(aes(x = median_income,
             y = median_house_value))+
  geom_point()+
  geom_smooth(method = "lm", se = FALSE)
```


```{r}
houses %>% 
  ggplot(aes(x = ocean_proximity,
             y = median_house_value))+
  geom_boxplot()
```


Shortly we may try a regression model to fit the categorical predictor ocean_proximity. Investigate the level of ocean_proximity predictors. How many dummy variables do you expect to get from it?

```{r}
houses %>% 
  distinct(ocean_proximity)
```


```{r}
alias(lm(median_house_value ~ ., data = houses))
```


```{r}
model_value_income <- lm(median_house_value ~ median_income,
                         data = houses)

autoplot(model_value_income)
```


```{r}
summary(model_value_income)
```


```{r}
model_value_income_ocean <- lm(median_house_value ~ median_income + ocean_proximity,
                               data = houses)

autoplot(model_value_income_ocean)
```


```{r}
summary(model_value_income_ocean)
```


